|

### 2021-08-04 appendix/appendix.Rmd
changes article and appendix - adding model 1

---
title: |
       | Supplementary Materials for "The Effect of Legislature Size on Public Spending: A Meta-Analysis"
author:
- "Huzeyfe Alptekin[^alptekin]"
- "Danilo Freire[^freire]"
- "Umberto Mignozzetti[^mignozzetti]"
- "Catarina Roman[^roman]"
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^alptekin]: Research Associate, Contemporary Brazilian History Research and Documentation Center, School of Social Sciences, Fundação Getulio Vargas, <huzeyfealptekin@gmail.com>.

[^freire]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

[^mignozzetti]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>.

[^roman]: PhD Student, University of California San Diego, <catarinamroman@gmail.com>, <http://catarinaroman.github.io>.

\appendix

@@

```{r, include=FALSE}
## Starting
set.seed(732578) # From random.org

# Required packages
pkgs <- c("tidyverse", "meta", "metafor",
          "readxl", "devtools", "data.table",
          "knitr", "gridGraphics", "gridExtra",
          "ggpubr", "kableExtra", "magick",
          'stargazer', 'pander', 'broom')

# Install if not already installed
install <- function(x) {
  if (x %in% rownames(installed.packages()) == FALSE)
    install.packages(x, dependencies = T,
                     repos = "http://cran.us.r-project.org")
}
lapply(pkgs, install)
devtools::install_github("isubirana/compareGroups")
devtools::install_github("MathiasHarrer/dmetar")

# Load packages
lapply(pkgs, require, character.only = T)
library("compareGroups"); library(dmetar)

# Broom the multi-level model
broom_mod <- function(mod, subgroup = FALSE) {
  aux <- data.frame(predict(mod))
  if (subgroup) {
    teeff = aux$pred
    seteeff = aux$se
    labeff = c("Subgroup Effect")
    loeff = c(aux$ci.lb)
    upeff = c(aux$ci.ub)
  } else {
    teeff = c(aux$pred, NA)
    seteeff = c(aux$se, NA)
    labeff = c("Overall Effect",
               "Prediction Interval")
    loeff = c(aux$ci.lb, aux$cr.lb)
    upeff = c(aux$ci.ub, aux$cr.ub)
  }
  mod2 <- tibble(
    TE = as.numeric(mod$yi),
    seTE = sqrt(mod$vi),
    studlab = mod$slab,
    lower = as.numeric(mod$yi)-1.96*sqrt(mod$vi),
    upper = as.numeric(mod$yi)+1.96*sqrt(mod$vi),
    group = 'A') %>%
    bind_rows(.,
              aux = tibble(
                TE = teeff,
                seTE = seteeff,
                studlab = labeff,
                lower = loeff,
                upper = upeff,
                group = "B")) %>%
    group_by(studlab) %>%
    mutate(studlab2 = paste0(studlab, "_", 1:n())) %>%
    ungroup()
  return(mod2)
}

# Estimation of the 
estim_het <- function(dat, yi, v, random, slab, hetvar = NULL) {
  if (is.null(hetvar)) {
    stop('Hetvar should be different than null.')
  }
  levshetvar <- unique(hetvar)
  fullmod <- rma.mv(yi = yi, V = v, random = random,
                    data = dat, slab = slab, method = 'REML',
                    test = 't', tdist = T)
  res <- tibble()
  for (i in levshetvar) {
    partmod <- rma.mv(yi = yi, V = v, 
                      random = random,
                      data = dat, slab = slab, 
                      method = 'REML',
                      test = 't', tdist = T, 
                      subset = hetvar == i)
    partmod <- broom_mod(partmod, subgroup = T)
    partmod$byvar = i
    partmod <- bind_rows(
      tibble(TE = NA, seTE = NA,
             studlab = toupper(i), lower = NA,
             upper = NA, group = "B", byvar = i),
      partmod
    )
    res <- bind_rows(res, partmod)
  }
  aux <- data.frame(predict(fullmod))
  res <- bind_rows(res,
                   tibble(
                     byvar = NA,
                     TE = c(aux$pred, NA),
                     seTE = c(aux$se, NA),
                     studlab = c("Overall Effect", "Prediction Interval"),
                     lower = c(aux$ci.lb, aux$cr.lb),
                     upper = c(aux$ci.ub, aux$cr.ub),
                     group = "B"))
  res <- data.frame(res)
  res$byvar <- toupper(res$byvar)
  return(res)
}

# Build plot function for forest plots
build_forest <- function(mod, capt, lsize = 22, ttl = NULL) {
  if(class(mod)[1] == "rma.mv") {
    mod2 <- broom_mod(mod)
  } else {
  # Build dataset for plot
  mod2 <- tibble(
    TE = mod$TE,
    seTE = mod$seTE,
    studlab = mod$studlab,
    lower = mod$lower,
    upper = mod$upper,
    group = "A") %>%
    bind_rows(.,
              aux = tibble(
                TE = c(mod$TE.random, NA),
                seTE = c(mod$seTE.random, NA),
                studlab = c("Overall Effect",
                            "Prediction Interval"),
                lower = c(mod$lower.random,
                          mod$lower.predict),
                upper = c(mod$upper.random,
                          mod$upper.predict),
                group = "B")) %>%
    group_by(studlab) %>%
    mutate(studlab2 = paste0(studlab, "_", 1:n())) %>%
    ungroup()
  }
  # Graph limits
  limg <- max(abs(c(mod2$lower, mod2$upper)))
  # Build plot
  p <- mod2 %>%
    ggplot(aes(y = reorder(studlab2, TE),
               x = TE, xmin = lower, xmax = upper)) +
    geom_point(aes(color = group)) +
    geom_errorbarh(aes(color = group),
                   height = 0.1) +
    scale_color_manual(values = c("#000000", "#8b0000")) +
    scale_x_continuous(limits = c(-1.1 * limg, 1.1 * limg)) +
    scale_y_discrete(
      labels = function(x)
        str_replace(x, "_[0-9]*$", "")) +
    geom_vline(xintercept = 0,
               color = "#000000", linetype = "dashed") +
    labs(x = "",
         y = "") +
    facet_grid(group~., scales = "free", space = "free") +
    labs(caption = capt,
         title = ttl) +
    theme_minimal() %+replace%
    theme(strip.text.y = element_blank(),
          legend.position = "none",
          axis.text.y = element_text(size = .8 * lsize,
                                     hjust = 1),
          axis.text.x = element_text(size = .6 * lsize,
                                     hjust = 1.1),
          plot.caption = element_text(size = lsize),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5,
                                    face = "bold",
                                    margin = margin(0, 0, 10, 0)),
          panel.grid.major = element_blank())
  return(p)
}

# Build forest plot for heterogeneous analysis
build_forest_het <- function(dat, coef, v, slab, capt, lsize = 22, ttl = NULL, hetvar = NULL) {
  mod2 <- estim_het(dat = dat, 
                    yi = coef, 
                    v = v, 
                    random = ~ 1 | id_level1/id_level2, 
                    slab = slab, 
                    hetvar = hetvar)
  mod2 <- data.frame(mod2)
  mod2$byvar <- toupper(mod2$byvar)
  TEaux <- mod2$TE
  TEaux[mod2$studlab== 'Subgroup Effect'] = TEaux[mod2$studlab== 'Subgroup Effect'] - 100
  # Graph limits
  limg <- max(abs(c(mod2$lower, mod2$upper)))
  # Build plot
  p <- mod2 %>%
    ggplot(aes(y = reorder(studlab, TEaux),
               x = TE, xmin = lower, xmax = upper)) +
    geom_point(aes(color = group)) +
    geom_errorbarh(aes(color = group),
                   height = 0.1) +
    scale_color_manual(values = c("#000000", "#8b0000")) +
    scale_x_continuous(limits = c(-1.1 * limg, 1.1 * limg)) +
    scale_y_discrete(
      labels = function(x)
        str_replace(x, "_[0-9]*$", "")) +
    geom_vline(xintercept = 0,
               color = "#000000", linetype = "dashed") +
    labs(x = "",
         y = "") +
    facet_grid(byvar~., scales = "free", space = "free") +
    labs(caption = capt,
         title = ttl) +
    theme_minimal() %+replace%
    theme(strip.text.y = element_blank(),
          legend.position = "none",
          axis.text.y = element_text(size = .8 * lsize,
                                     hjust = 1),
          axis.text.x = element_text(size = .6 * lsize,
                                     hjust = 1.1),
          plot.caption = element_text(size = lsize),
          plot.title.position = "plot",
          plot.title = element_text(hjust = 0.5,
                                    face = "bold",
                                    margin = margin(0, 0, 10, 0)),
          panel.grid.major = element_blank())
  return(p)
}

# Load datasets
load("../dataset/dataCoefs.RData")
```

@@

# A Brief Introduction to the Law of _1/n_ and the Reverse Law of _1/n_

In this section, we present the intuition behind the _law of 1/n_, as well as the alternative _reverse law of 1/n_ proposed by @primo2008distributive.

In their seminal paper, @weingast1981political argue that a high number of legislators will increase public spending beyond the optimal economic benchmark. They suggest that politicians have an incentive to over-provide concentrated benefits to their constituencies, spreading the costs across all constituencies through generalised taxation. The corollary of their model is that larger legislatures generate more public spending.

In their model, every local public goods project of size $x$ generates a concave benefit $b(x)$, and there are convex costs associated with the project.\footnote{The concavity assumption implies that $b'>0$ and $b''<0$. A similarly convex assumption would mean that $c'>0$ and $c''>0$. We assume that the derivatives are well defined throughout the analysis. Moreover, we drop the constituency index to make the model more intuitive, so the reader should assume that we always employ a symmetric Nash equilibrium.} The first type of cost, $c_1(x)$, comprises the expenses within the constituency (e. g., hiring a local company for the project). The second type of cost, $c_2(x)$, captures the expenses outside the constituency (e. g., hiring a company from another state). Finally, the third cost, $c_3(x)$, captures the externalities generated by the project (e. g., how much prices shift because local economic factors are being used to provide the project). The total cost is equal to $c(x) = c_1(x) + c_2(x) + c_3(x)$. The tax burden generated by the project is equal to $T(x) = c_1(x) + c_2(x)$.

Projects are economically efficient when the marginal costs are equal to the marginal benefits of the project size. This leads to the economic optimal project size, $x^E$. The optimal project size is $b'(x^E) - c_1'(x^E) - c_2'(x^E) - c_3'(x^E) = 0$. However, the projects that are actually implemented have a different structure. First, assume that the constituency in question has a tax burden $t = 1/n$, where $n$ represents the number of constituencies. Also, suppose that benefits are distorted by the fact that costs within the constituency ($c_1(x)$) become investments in local firms. Therefore, the costs and benefits of implementing a project with size $x$ have the following structure:

\[
N(x) = b(x) + c_1(x) - \frac{1}{n}[c_1(x) + c_2(x)] - c_3(x)
\]

Differentiating $N(x)$ with respect to $x$ gives us the first order condition for an optimal project implementation.

\[
b'(x) + c_1'(x) - \dfrac{1}{n}[c_1'(x) + c_2'(x)] - c_3'(x) = 0
\]

Totally differentiating $x$ with respect to $n$ gives us the following:

\[
b''(x)\dfrac{dx}{dn} + c_1''(x)\dfrac{dx}{dn} - \dfrac{1}{n}[c_1''(x) + c_2''(x)]\dfrac{dx}{dn} + \dfrac{1}{n^2}[c_1'(x) + c_2'(x)] - c_3''(x)\dfrac{dx}{dn} = 0
\]

And rearranging the terms, we find the following expression for $\dfrac{dx}{dn}$:

\[
\dfrac{dx}{dn} = -\dfrac{n^{-2}[c_1'(x) + c_2'(x)]}{b''(x) + c_1''(x)-n^{-1}[c_1''(x) + c_2''(x)] - c_3''(x)}
\]

Note that the numerator is always positive, as the marginal investment costs inside and outside the district increase according to the project size. The _law of 1/n_ holds when the denominator is negative. $b''(x) < 0$ by assumption, and $-n^{-1}[c_1''(x) + c_2''(x)] - c_3''(x) < 0$ as $c_1'', c_2'', c_3'' > 0$. Then, the _law of 1/n_ is true when $c_1''(x)$ is smaller than $n^{-1}[c_1''(x) + c_2''(x)] + c_3''(x) - b''(x)$.

As the above condition suggests, the _law of 1/n_ holds only in specific situations. @primo2008distributive advance that theory by considering other situations where the _law_ may not hold, and they also argue that there are cases in which a _reverse law of 1/n_ holds, that is, larger legislatures may lead to lower public expenditures.

Following @primo2008distributive, let $n$ be the number of districts, $m$ the number of citizens in each district, and $nm$ is the total population in the country. Consider a local public good that generates a per capita benefit of $b(x, m) = x^\alpha m^{\beta-1}$ according to size $x$, where $\beta$ is the degree of congestion of the public good, that is, how much the addition of individuals reduce the benefits for other individuals (note that the lower the $\beta$, the higher the congestion). In terms of costs, consider a linear cost function $C(x) = x$, and in terms of taxation, assume that the people in the district pay both local and federal government taxes. The degree that taxes are shared is denoted by $s$. Moreover, there is a deadweight loss of the taxes $\theta \geq 1$. The tax then becomes: $t = \left(\dfrac{(n – ns + s)x + (ns – s)X}{nm}\right)^\theta$. Therefore, the citizens receive the following net benefit of a project with size $x$:

\[
\pi = x^\alpha m^{\beta-1} - \left(\dfrac{(n – ns + s)x + (ns – s)X}{nm}\right)^\theta
\]

Maximising this function, and solving for the symmetric Nash equilibrium where $x = X$ for all projects, we find the following optimal project size:

\[
x^* = \left[\left(\dfrac{\alpha}{\theta}\right)^{\dfrac{1}{\theta-\alpha}}\right]\left[\left(nm\right)^{\dfrac{\beta+\theta-1}{\theta - \alpha}}\right]\left[\left( \dfrac{n^{2-\beta-\theta}}{n-ns+s}\right)^{\dfrac{1}{\theta-\alpha}}\right]
\]

And in the graph below, we show simulations for $n$ varying from 10 to 20, holding constant $\beta = 0.35$, $\theta \in \{0.65, 0.75\}$, $s = 0.5$, $m = 100$, and $\alpha = 0.7$.

```{r, echo=FALSE, cache=TRUE, fig.align='center'}
xstar <- function(x, beta, theta, alpha = 0.7, s = 0.5, m = 10) {
  n = x
  a = (alpha/theta)^(1/(theta-alpha))
  b = (n*m)^((beta+theta-1)/(theta-alpha))
  c = ((n^(2-beta-theta))/(n-n*s+s))^(1/(theta-alpha))
  return (a*b*c)
}

curve(xstar(x, beta = 0.35, theta = 0.8), from = 10, to = 20,
      main = 'Increasing project size when varying legislature size \n (high deadweight losses)', xlab = 'Legislature Size', ylab = 'Project Size')
curve(xstar(x, beta = 0.35, theta = 0.5), from = 10, to = 20,
      main = 'Decreasing project size when varying legislature size \n (low deadweight losses)', xlab = 'Legislature Size', ylab = 'Project Size')
```

Thus, we see that both the _law of 1/n_ and its reverse formulation are equally plausible. Here we show how the results change just by changing the levels of deadweight losses, but authors have suggested other reasons why the _law of 1/n_ may not apply, such as bicameralism [@chen2007law], popular initiatives [@matsusaka2005endogeneity], type of government [@coate2011government], supermajorities [@lee2015supermajority], political fragmentation [@lledo2003electoral], and ideology [@bjedov2014impact].

# Search Criteria

The first step in our systematic review consisted in gathering a study sample. We started our data collection with a manual search based on a set of keywords we scouted from the distributive politics literature. This search produced a database with many entries that were unrelated to our subject of investigation. To reduce the number of false positives in our sample, we restricted our search to studies that cited Weingast, Shepsle and Johnsen's 1981 paper "_The Political Economy of Benefits and Costs: A Neoclassical Approach to Distributive Politics_", which is the fundamental contribution to the field. Although \href{https://scholar.google.com/}{Google Scholar} reports the article has received \href{https://scholar.google.com/scholar?um=1&ie=UTF-8&lr&cites=13117579863846712459}{2,180} citations, our search resulted in 2,664 records on the 21^st^ of November 2019.

We webscraped three large academic databases: \href{https://scholar.google.com/}{Google Scholar} (n = 1001); \href{https://academic.microsoft.com/home}{Microsoft Academic} (n = 927); and \href{https://www.scopus.com/}{Scopus} (n = 736). The `R` script we wrote extracted the article title, abstract, authors, year, journal of publication, and database from which the record originated. Our code is available in the `R` script available in this repository. We screened these results with an English language and article restriction, that is, we excluded all records written in other languages and all that were not academic papers, such as book chapters or doctoral theses. We set no restriction to unpublished articles.

We complemented this process by doing a term search on Google Scholar after having finished the entire eligibility procedure in the first sample. We formulated the search string based on the terms and expressions that appeared most frequently in the articles included in our meta-analysis. The choice of string translates the central point of our inquiry: the relationship between legislature size and public spending. The search string was as follows: `("upper chamber size" OR "lower chamber size" OR "council size" OR "parliament size" OR "legislature size" OR "number of legislators" OR "legislative size") AND ("spending" OR "expenditure" OR "government size")`. We scraped Google Scholar on March 5^th^, 2021. Using the titles of this new database's records, we did a fuzzy matching with the database of @weingast1981political citations mentioned previously. We then performed the exact same eligibility procedures, further explained below. This resulted in 3,041 additional records. Combining the two search strategies, we assessed a total of 5,705 records.

# Article Selection

The selection process was conducted by two authors in three phases. In the first phase, we excluded all titles that were clearly unrelated to our topic of interest. This was only a preliminary step, as we were not able to eliminate a large number of entries. Then, we read all abstracts. We kept all publications whose main topics were either government expenditure or legislative structures. Abstracts that indicated that the paper discussed or estimated the impacts of representative institutions, elections, or chamber dynamics were included. This allowed us to significantly reduce our sample.

In the second phase, we assessed full texts. To remain in our sample, the paper should (i) conduct a quantitative analysis, (ii) report data on the number of legislators, and (iii) also include data on public expenditure. If the publication had all three, we kept it in our sample. Disagreements in this phase were discussed among the authors, and a third investigator was consulted when needed.

The third phase consisted of filling out tables for each of the remaining 47
articles to systematically evaluate their eligibility. Since authors use
different measures for government spending and the number of lower/upper house
members, we extracted all coefficients that provided this information. In this
phase, we also collected information on whether the paper had been published,
and if it explicitly discussed the _law of 1/n_. Upon choosing the variables,
we excluded two studies from the first sample, as they did not have the
dependent/independent variables we collected for our meta-analysis. We then
included the 2 papers we found in the second search, as both conformed to our
criteria, and compiled our final sample of `r length(unique(dat$authoryear))`
articles.

## Exclusion Analysis

We selected the final pool of articles based on two criteria, namely the independent and the dependent variables employed in the paper. The categories follow below:

1.  Independent variables:

-   *N*: Number Legislators in the Lower Chamber
-   *logN*: Log Number Legislators in the Lower Chamber
-   *K*: Number Legislators in the Upper Chamber

2.  Dependent variables:

-   *ExpPC*: Expenditure Per Capita
-   *logExpPC*: Log Expenditure Per Capita
-   *PCTGDP*: Expenditure as Percentage of GDP

If a paper did not use a combination of these variables, we excluded it from the meta-analysis.

## Flow Chart

The diagram below shows each step of our article selection process. We followed the Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) statement to conduct our study\footnote{More information about the PRISMA statement is available at \url{http://www.prisma-statement.org}.}. The column to the right depicts the amount of articles excluded in each phase, and the one to the left shows the number of records evaluated.

\bigskip

```{tikz tikz-ex, echo=FALSE, cache=TRUE, eval=TRUE, engine.opts = list(template = "tikz2pdf.tex"), fig.cap = "PRISMA Flow Chart for First Search Strategy: Weingast et al. Citations"}
\usetikzlibrary{shapes.arrows,arrows.meta,positioning,shapes.geometric}
  \begin{tikzpicture} [node distance = 2.7cm, xshift = 1.5]
  \footnotesize
  \linespread{1.0}

  \node (ID) [stage, rotate = 90, yshift = 2cm, xshift = 2.5cm] {\normalsize{Identification}};

  \node (SC) [stage, below of = ID, rotate = 90] {\normalsize{Screening}};

  \node (EL) [stage, below of = SC, rotate = 90] {\normalsize{Eligibility}};

  \node (INC) [stage2, below of = EL, rotate = 90, xshift = -1.2cm] {\normalsize{Included}};

  \node (id) [phase, right of = ID, xshift = 1cm] {Records identified through webscraping  \\ \textbf{(n = 2664)}};

  \node (1st) [exc, right of = id, xshift = 3cm] {Records in languages other than English, book chapters, doctoral theses, and duplicates excluded \\ \textbf{(n = 1220)}};

  \node (screen) [phase, below of = id] {Records screened \\ \textbf{(n = 1445)}};

  \node (2nd) [exc, right of = screen, xshift = 3cm] {Records excluded after reading title and abstract \\ \textbf{(n = 1069)}};

  \node (elig) [phase, below of = screen] {Full-text articles assessed for Eligibility \\ \textbf{(n = 376)}};

  \node (3rd) [exc, right of = elig, xshift = 3cm] {Non-quantitative studies or records using unrelated variables excluded \\ \textbf{(n = 329)}};

  \node (inc_1) [phase, below of = elig] {Preliminary included articles \\ \textbf{(n = 47)}};

  \node (4th) [exc, right of = inc_1, xshift = 3cm] {Articles excluded during analysis due to nonconforming variables \\ \textbf{(n = 19)}};

\node (inc_2) [phase, below of = inc_1] {Included articles \\ \textbf{(n = 28)}};
  \draw [arrow] (id) -- (1st);
  \draw [arrow] (id) -- (screen);
  \draw [arrow] (screen) -- (2nd);
  \draw [arrow] (screen) -- (elig);
  \draw [arrow] (elig) -- (3rd);
  \draw [arrow] (elig) -- (inc_1);
  \draw [arrow] (inc_1) -- (4th);
  \draw [arrow] (inc_1) -- (inc_2);

 \end{tikzpicture}
```

\bigskip

>>>>

---
title: |
       | Supplementary Materials for "Legislature Size and Welfare: Evidence from Brazil"
author:
  - Umberto Mignozzetti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>. Corresponding author.

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

[^gabriel]: Associate Professor, Department of International Relations, São Paulo State University, <gabi.cepal@gmail.com>.

\appendix
\addtocounter{section}{2}

@@

```{r, include=FALSE}
## Starting
XXXX

# Required packages
XXXX

# Load datasets
XXXX
```

@@

\subsection{A Model of Legislature Size and Service Provision}

Consider an strategic interaction comprised of a mayor ($M$), and $N \geq 3$ of city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is head of the local Executive, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. The policy proposal, if accepted, is implemented.

Every city councilor has a party affiliation, and the party label may or may not match the mayor's affiliation. If the politician shares the same party as the mayor, we say that it belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, we assume that the politician belongs to the opposition $O \not\subset G$. The party affiliation is important because it provides a political benefit or cost for the councilor. If the mayor implements the policy, then a councilor aligned with the mayor receives a policy benefit $p > 0$. If the politician belongs to the opposition, then the mayoral policies impose her a cost of $-p$. For simplicity, we assume that these costs are symmetric.

However, if the mayor's policy proposal lose majoritarian support from the council, then we assume that a reversal policy takes place, where the resources diminish by a factor of $\delta \in (0,1)$, and a random councilor is recognized to make a proposal.\footnote{In a model similar to [Baron Ferejohn / Cliff XXXX].} The councilor proposes the new policy vector, and the council votes. If the policy is rejected, then another councilor is recognized to make the offer, and this goes on until one offer is finally accepted by the majority of the legislators. For simplicity, we assume that the councilor's offers generate no political benefits or costs, so that if a councilor proposes the policy vector, then $p=0$. This may sound restrictive, but we will show that councilors propose zero public goods in equilibrium, as they have only to please their particularistic constituencies to get reelected.

To convince the councilors, the politician making the proposal has three vectors of policies at her disposal. First, the proposer may give targeted goods for the councilor $i$ that she is trying to earn support, $t_i \geq 0$. These targeted goods improve the reelection rates of the councilor. Second, the proposer may also give private goods $x_i \geq 0$ to a councilor $i$. Private goods do not affect the reelection rates of the politicians, but provide instant benefits for the receiver. Finally, the proposer can provide public goods $g \geq 0$. Public goods help the mayor to get reelected, and also affect the policy benefits received by the politicians, depending on the party affiliation.

For the mayor, the expected utility function is a sum of the utility from rents and benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of getting reelected $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of getting reelected is also multiplied by the benefit from holding office $B_M > 0$. This term captures the tangible and intangible gains that the mayor perceives from being reelected into the position. The utility is multiplied by an indicator $\mathbb{I}$ that captures whether the mayor's proposal was accepted by the city council.\footnote{The indicator function is defined as
\[
I(condition) \ = \ \twopdo{1}{\text{Condition is satisfied.}}{0}.
\]
} The expected utility for the mayor is:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \ \mathbb{I}(\text{Approval})(u(r) + B_M\pi(g))
\]

And the mayor chooses the optimal policy subject to a budget constraint. The municipality has $R$ resources, and we assume that the municipality cannot run debts. This budget balance constraint requires the policy vector $(r, g, \bf{x}, \bf{t})$ to satisfy the following rule:

\[
r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ \delta^kR
\]

And the mayor's proposal happens when $k=0$. This makes mayoral budget inequality to satisfy: $r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ R$. Otherwise, in the reversal proposals, the resources diminish by a $\delta^k$ rate, with $k$ being the number of unsuccessful offers by the councilors.

We can decompose the councilor's utility function into three major components. First, the political gains or losses (depending on the councilor's party affiliation); second, the rents $x_i$ offered by the policy proposer to the councilor; and finally, the transfers for the constituents $t_i$, offered by the policy proposer to the councilor. As the constituent transfer $t_i$ affects chance of winning the election $\pi_C(t_i)$, we multiply this chance by the benefits accrued from remaining in office, $B > 0$. Again, as in the mayor's case, the utility is only realized when the city council accepts the policy proposed, so that we multiply the approval indicator $\mathbb{I}(.)$. The councilor's expected utility is equal to:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \  \mathbb{I}(\text{Approval})((-1)^{\mathbb{I}(i \in O)}p + x_i + B\pi_C(t_i))
\]

For simplicity, we assume that the councilor's probability of reelection satisfy a simple cutoff rule:

\[
\pi_C(t_i) \ = \ \twopdo{1}{t_i \geq \underline{t}}{0}
\]

This assumption simplifies the calculations, and it is reasonable as many models have the voter's decision as a cutoff strategy [@XXXX]. We also assume, for convenience, that $B > \underline{t}$. Otherwise, the councilor would prefer to pocket the targeted goods, and not being reelected.

The timeline of the game is as follows:

1. The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
2. The mayor proposes a policy vector $(r, g, \bf{x}, \bf{t})_M$.
3. The city council votes the proposal.
  - If the council accepts, then the policy is implemented, and the game ends. 
  - Otherwise, the reversal policy is implemented.
4. (Reversal Policy) At the beginning of each iteration of the reversal policy, one councilor is randomly recognized to be the proposer. She proposes a vector $(\bf{x}, \bf{t})$, with $r =  g = 0$. If the proposal is accepted, it is implemented. If the proposal is rejected, a new round of reversal start, but with the municipal resources reduced by a factor $\delta$.

We solve the game for a Sub-game Perfect Nash Equilibrium using backward induction. Although the game tree may extend infinitely, a Stationary Sub-game Perfect Equilibrium can be reached with no delay in the policy implementation, and the mayoral proposal being successful.

\begin{definition}
The solution for the game consists in a pair of XXXX.
\end{definition}

Solving by backward induction, suppose that there were $k$ prior rejections, and the municipal resources shrank by $\delta^{k}R$. As at this stage the politicians are similar with regard to political affiliation, we can drop the party label.\footnote{In the extensions, we solve the game assuming that the distribution remained partisan even for the councilors. The results are similar, with the only difference that when partisanship is relevant for councilors, the expected provision of public goods is higher than zero.} A given councilor is recognized as the proposer with $1/N$ probability. The recognized councilor has to convince at least one-half of the council members to agree with her policy proposal. Therefore, a legislator $i$, that receives an offer from a proposer ($x_i, t_i$), at round $k$, will accept the offer if it is at least as good as waiting for the next round.

\[
x_i + B\pi_C(t_i) \ \geq \ \dfrac{1}{N}\left[\delta^{k+1}R - \dfrac{N}{2}(x_{-i} + t_{-i}) + B\pi_C(t_i)\right] + \dfrac{1}{2}(x_i + B\pi_C(t_i))
\]

At the $k$-th stage, the players are equivalent. Therefore, the offer $x_i, t_i$, and the offer for the other players, ($x_{-i} + t_{-i}$), are the same. Moreover, as $B > \underline{t}$, $t^* = t_i = t_{-i} = \underline{t}$, and the offer that would make legislator $i$ accept without further delay is:

\[
x + B \ \geq \ \dfrac{1}{N}\left(\delta^{k+1}R - \dfrac{N}{2}(x + \underline{t}) + B\right) + \dfrac{1}{2}(x + B)
\]

And after some algebra, the optimal private goods offer has to satisfy:

\[
x \geq \dfrac{1}{N}\left( \delta^{k+1}R + B \right) - \dfrac{1}{2}(B + \underline{t}) \equiv \underline{x}(k)
\]

For all $k \in \{1, 2, \cdots, \infty\}$.

\begin{lemma}
Suppose that the council rejected the mayor's offer. Then, at the $k$-th stage of the game, the proposer will offer $t^* = \underline{t}$ and $x^* = \underline{x}(k)$ for half of the members, keeping for her $t_i = \max\{0, \underline{t}\}$, and

\[
x_i \ = \ R\delta^k\left( 1 - \dfrac{\delta}{2} \right) + \dfrac{N}{4}(B-\underline{t}) + \dfrac{B}{2}
\]

Subject to the budget constraint:

\[
\dfrac{N}{2}\left(\underline{t}+\underline{x}(k)\right) + x_i  \ \leq \ \delta^kR
\]

The sequence of strategies derived here are an equilibrium for all $k \in \{1, 2, \cdots, \infty\}$.

\end{lemma}

\begin{proof}
The proof is direct, after collecting the terms from the previous discussion.

Note that after some iterations, the budget may shrink enough so that $t^* = \underline{t}$ is unfeasible. This means that there exists a $k^*$ such that, when $k > k^*$, $t^* = 0$, and the councilors divide the remaining budget as rents. This case is trivial to solve, making $\underline{x}(k) = \dfrac{1}{N}\left(\delta^{k+1}R\right)$ and $x_i = \delta^kR\left(1 - \dfrac{\delta}{2}\right)$.
\end{proof}

We now have the entire strategy profile for the city councilors, provided that they reject the initial mayor's proposal. Proceeding by backward induction, we need to find the mayor's proposal that will win the support of the majority of the council. As the mayor is the proposer, policy concerns come into play, and the offer for a government aligned councilor will be different from the offer for an opposition councilor. For a government aligned councilor, assuming that the councilor is the pivotal players, the accepted offer has to be better than rejecting the proposal and waiting for the next round:

\[
p + x_i + B\pi_C(t_i) \ \geq \ \dfrac{1}{N}\left(\right) + \dfrac{1}{2}(x_i + B\pi_C(t_i))
\]

.
